Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-10-11 paper/sections/proof-reconstruction.tex
improved some references in the text.

\begin{figure}
\centering

\begin{tikzpicture}[scale=0.9]
\node[text width=2cm, align=center](problem) at (0,0)
  {1. \CPL \\ Problem.};

\node[right = 1.1cm of problem, text width=2cm, align=center]
  (tptp){2. \TPTP \\ Problem.};

\node[right= 1.1cm of tptp, text width=2cm, align=center]
   (metis) {3. \Metis \\ Prover.};

\node[right= 1.1cm of metis, text width=2cm, align=center, inner sep=10pt]
   (tstp) {4. \TSTP \\ Derivation.};

\node[below= 0.5cm of tstp, text width=2cm, align=center, inner sep=10pt]
  (athena) {5. \Haskell \\ Translator.};

\node[left = 1.1cm of athena, text width=2cm, align=center]
   (agdafile) {6. Proof-term.};

\node[left = 1.1cm of agdafile, text width=2cm, align=center]
   (agda) {7. Proof Checker.};

\node[below = 0.5cm of problem, text width=2cm, align=center]
   (verified) {8.1 Proof \\ Checked.};

\node[below = 0.5cm of verified, text width=2cm, align=center]
   (failure) {8.2 Failure.};

% node[below] {send to}
\draw[->, thick] (problem) to
  % node[below] {\tiny encoding}
  (tptp);
\draw[->, thick] (tptp) to
  % node[below] {\tiny }
  (metis);
\draw[->, thick] (metis) to
  % node[below] {\tiny replies on}
  (tstp);
\draw[->, thick] (tstp) to
  % node[right] {\tiny parsing}
  (athena);
\draw[->, thick] (athena) to
  % node[below] {\tiny traslation}
  (agdafile);
\draw[->, thick] (agdafile) to
  % node[below] {\tiny type-checking}
  (agda);
\draw[->, thick] (agda) to (verified);
\draw[->, thick, gray] (agda) to (failure);
\end{tikzpicture}
\caption{Proof-reconstruction overview.}
\label{fig:proof-reconstruction-workflow}
\end{figure}

The proof-reconstruction approach proposed here consists of a series
of steps similar to the workflow presented by \citeauthor{sultana2015} in
\cite{sultana2015}. This process is a translation from a source
system to a target system. In our case, the system of origin, the
automatic theorem prover, is \Metis; the target system is a proof
assistant, \Agda. We choose \Agda, but another proof assistant with the same support of type theory and inductive types is valid (see a description of \Agda for its main features in
Section~\ref{sec:implementation}).
The overview of the proof-reconstruction is presented in
Fig.~\ref{fig:proof-reconstruction-workflow}, the process can begin with
a problem in \CPL that we encode in \TPTP format to be sent as the
input of the \Metis prover.
The algorithms of \Metis in Step~3 find
a solution if the problem is a theorem, in that case \Metis replies
with a derivation in \TSTP format of the proof, otherwise the problem is counter feasible.

With such a derivation, in Step~5, we process the derivation with our \Haskell translator tool, \Athena~\cite{Athena}.
\Athena parses the \TSTP format, analyze the
derivation and generates a representation of the natural deduction
proof using a tree data structure (see the properties
of this tree in Subsection \ref{ssec:metis-proofs}).
We perform an analysis of the proof-tree removing redundancies,
unnecessary steps or unused output from the \TSTP derivation.

As a result,
\Athena generates an \Agda file with the proof-term. We have included
in this file, \emph{imports} that call our \Agda libraries that give
support for CPL presented in~\cite{AgdaProp}, and support for \Metis'
inference rules implemented in~\cite{AgdaMetis}.

Finally, we use the proof checker in Step~7 to type-check the
proof-term. If type-checking succeed, the \TSTP
derivation is correct, otherwise when it fails, one of the three
actors in the process could be the responsible: \Metis, \Athena, or
\Agda.

We are responsible for errors committed in the translation from
\TSTP to \Agda: bugs in the parsing module, lack of the coverage of
\Metis' inference rules, and last but not least, the pretty print of
the proof in \Agda code.

In the rest of this section, we provide a formal description using
type theory to build definitions and theorems for functions
necessary to emulate \Metis' inference rules.
Some conventions about notations. We will write the
type judgments with ``$a : A$'' to denote that $a$ has type
$\abbre{A}$. Types are written using small caps.
For propositions, we will denote by $φ ≡ ψ$, the syntactic
equivalence. We use the \Prop type for propositions and for theorems,
the dependent type, $ Γ ⊢ φ$ where $Γ$ has type $\abbre{List} \Prop$ and
$φ$ has type $\Prop$.

% ...................................................................

% section section_name (end)
\subsection{Structural Recursion}
\label{ssec:structural-recursion}

To emulate \Metis' inference rules in type theory,
we observed that transformations performed by some rules were given by \emph{general recursive} functions.
But algorithms based on general recursions can not be translated
directly in type theory since it is not a guarantee they terminate.

Then, we solve the termination problem by modifying problematic functions
to be \emph{structural recursive}~\cite{Coquand1992,Abel2002,Bove2005}.

We use the same notion of \emph{structurally smaller} from
\citeauthor{Abel2002}~\cite{Abel2002} to say, a recursive function is
structurally recursive if it calls itself with only
\emph{structurally smaller}
arguments.

To convert a general recursive call into a structurally recursive call,
we use the \emph{Bounded Recursion} technique presented
in~\cite{Bertot2004} but
others found in \cite{Coquand1992,Abel2002,Bove2005} could be used.
This technique converts a general recursive
function into a structural recursive function by adding an additional
argument for the function complexity, the bound for the number of times, the function needs to call itself to get the final result.

Then, we add the natural number $n : \Nat$ in the input of the
general recursive function, and perform the recursion over the structure of
\Nat type. \abbre{Nat} type is the usual type for natural numbers with \zero and \suc as their constructors. We consider the common arithmetic operations ($+,\,-,\,*$) and numbers $0,1,2,\dots$ as representations of $\zero, \suc, \suc(\zero), \dots$.

Then, the type-checker will
recognize structurally smaller arguments by using the lexicographic
order~\cite{Abel2002} for such kind of functions.

In other words, we take a general recursive function $f: \Prop \to \Prop$.
and we redefine that function as
$f: \Prop \to \Nat \to \Prop$, and perform the recursion
using induction on the structure of the \Nat type.

Let us consider the following example to exhibit the technique
mentioned above for an inner function in the \strip inference.

\begin{align}
\begin{split}
\label{eq:unshunt-definition}
&unshunt_{0}(φ) =\\
&\hspace{3mm}\begin{cases}
unshunt_{0}((φ₁ ∧ φ₂) ⇒ φ₃),
  & \text{if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
unshunt_{0}(φ₁ ⇒ φ₂) ∧ unshunt_{0}(φ₁ ⇒ φ₃),
  & \text{if }φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
φ, &\text{otherwise.}
\end{cases}
\end{split}
\end{align}

From the definition above, the size of the argument, ((φ₁ ∧ φ₂) ⇒ φ₃),
is not structurally smaller than the input formula, (φ₁ ⇒ (φ₂ ⇒ φ₃)), since ((φ₁ ∧ φ₂) ⇒ φ₃) is not a subformula of (φ₁ ⇒ (φ₂ ⇒ φ₃)).

Thus, we define the function
$unshunt_{1} : \Prop \to \Nat \to \Prop$
by structural recursion in its second argument in
Eq.~\ref{eq:unshunt-structured}.

\begin{align}
\label{eq:unshunt-structured}
\begin{split}
&unshunt_{1}(φ, \zero)\hspace{3mm}= φ\\
&unshunt_{1}(φ, \suc(n)) =\\
&\hspace{3mm}\begin{cases}
unshunt_{1}((φ₁ ∧ φ₂) ⇒ φ₃, n),
  &\text{if } φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
unshunt_{1}(φ₁ ⇒ φ₂, n) ∧\ unshunt_{1}(φ₁ ⇒ φ₃, n),
   &\text{if } φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃).
\end{cases}
\end{split}
\end{align}

As we mentioned above, the natural number in the second argument of the
function used in Eq.~\ref{eq:unshunt-structured} stands for a complexity
measure of the function, in this case $unshunt_{1}$ function.

A \emph{complexity measure} defined in \cite{Agudelo-Agudelo2017}
and presented in \cite{VanDalen1994} as \emph{the rank for a
proposition}, is a function $\Prop \to \Nat$ that assigns a natural
number for a proposition φ as follows.

\begin{equation*}
\label{eq:rank-definition}
rank(φ)=
\begin{cases}
0, &\text{for atomic }φ \\
rank(φ₁) + c_{¬},  &\text{if } φ ≡ \neg φ₁ \\
\max{(rank(φ₁),rank(φ₂))} + c_{\square},
  &\text{if } φ ≡ φ₁\,\square\, φ₂.
\end{cases}
\end{equation*}

where $c_{¬}$ or $c_{\square}$ is a non-negative integer called the
complexity of the connective.

Recall, we want to maintain the same outputs as the original unshunt
does in Eq.~\ref{eq:unshunt-definition} and the complexity measure of it,
is a bound for the number of times this function can call itself.

Therefore, we define an appropriate complexity measure
$unshunt_{cm}: \Prop → \NN$ for our $unshunt_{1}$ function presented in
Eq.~\ref{eq:unshunt-rank-definition}.

\begin{align}
\begin{split}
\label{eq:unshunt-rank-definition}
&unshunt_{cm}(φ) =\\
&\hspace{3mm}
\begin{cases}
unshunt_{cm}(φ₃) + 2, &\text{if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
\max(unshunt_{cm}(φ₂),unshunt_{cm}(φ₃)) + 1,
  &\text{if }  φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
0, &\text{otherwise.}
\end{cases}
\end{split}
\end{align}

Finally, the curated version of $unshunt_{0}$ function is presented
in Eq.~\ref{eq:unshunt-fixed}.

\begin{equation}
\label{eq:unshunt-fixed}
unshunt(\varphi) := unshunt_{1}(\varphi, unshunt_{cm}(\varphi)).
\end{equation}

For a more complete treatment of general recursion, and how to
define well-founded and well-structural recursive functions, we refer
the reader to Ref.~\cite{Bove2005}, where Bove and Capettra described
not only general recursion, but they also treat mutual recursive
function, among other kinds.

% ................................................

\subsection{Emulating \Metis' Inference Rules}
\label{ssec:emulating-inferences}

\subsubsection{Splitting a Goal.}
The goal of a problem to be proved by \Metis is processed to generate new
propositions, the new targets, \Metis calls these formulas
\emph{subgoals}.

The conjunction of these subgoals is a proposition equivalent to the
original goal. Then, we find a refutation for each subgoal in the \Metis' \TSTP derivation.

Nevertheless, \Metis does not provide neither an explanation for the user about the process in its \TSTP derivations nor in its documentation.

In the following, we show how Theorem.~\ref{thm:thm-strip} proves the correctness of the process mentioned above. But first, let prove some lemmas for $unshunt$ function presented in Eq.~\ref{eq:unshunt-fixed},
and $split$ function from the definition in Eq.~\ref{eq:split-definition}.

\begin{lemma}[lem-unshunt]
  \label{lem:unshunt}
  $Γ ⊢ unshunt(φ)$, then $Γ ⊢ φ$.
\end{lemma}

\begin{proof} Use induction on the cases defined by the outcome of the
$unshunt$ function in Eq.~\ref{eq:unshunt-structured}.
When $n = 0$, by definition, $Γ ⊢ φ$.
When $n = 1$, we apply the following theorem from~\cite{AgdaProp}

\begin{equation*}
\texttt{∧⇒-to-⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
\end{equation*}

For $n > 1$, we will use induction on the structure of second
argument.

(φ₁ ⇒ (φ₂ ⇒ φ₃))
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
  \RightLabel{By def.}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
  \RightLabel{∧⇒-to-⇒⇒}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).}
  \end{bprooftree}
\end{equation*}

Second, using the following theorem from \cite{AgdaProp},

\begin{equation*}
  \texttt{⇒∧-to-⇒∧}\ :\ Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃) → Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃).
\end{equation*}

($φ₁ ⇒ (φ₂ ∧ φ₃)$)
\begin{equation*}
\scalebox{0.95}{
($\mathcal{D}_1$)
\begin{bprooftree}
\AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n) ∧ unshunt(φ₁ ⇒ φ₃, n)$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n-1)$}
\RightLabel{lem-unshunt}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$,}
\end{bprooftree}
}
\end{equation*}

\begin{equation*}
(\mathcal{D}_2)
\begin{bprooftree}
\AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂) ∧ unshunt(φ₁ ⇒ φ₃)$}
\RightLabel{∧-proj₂}
\UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₃, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₃, n-1)$}
\RightLabel{lem-unshunt}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$,}
\end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}_1$}
  \AxiomC{$\mathcal{D}_2$}
  \RightLabel{∧-intro}
  \BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
  \RightLabel{⇒∧⇒-to-⇒∧}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃).}
  \end{bprooftree}
\end{equation*}
\end{proof}

Now, let us consider the $split$ function presented in
Eq.~\ref{eq:split-definition} that yields a conjunction of subgoals from a proposition, the goal in \Metis.

\begin{equation}
\label{eq:split-definition}
\mysplit(φ) =
\begin{cases}
unshunt(\mysplit(φ₁)) &\text{if }φ ≡ φ₁ ∧ φ₂\\
∧\ unshunt(φ₁ ⇒ \mysplit(φ₂)), &\\[2.5mm]
unshunt(¬ φ₁ ⇒ \mysplit(φ₂)),
  & \text{if }φ ≡ φ₁ ∨ φ₂\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(φ₂)),
  & \text{if }φ ≡ φ₁ ⇒ φ₂\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(φ₂))
  & \text{if }φ ≡ φ₁ ⇔ φ₂\\
∧\ unshunt(φ₂ ⇒ \mysplit(φ₁)),
  &\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
  & \text{if }φ ≡ ¬ (φ₁ ∧ φ₂)\\[2.5mm]
unshunt(\mysplit(¬ φ₁))
  & \text{if }φ ≡ ¬ (φ₁ ∨ φ₂)\\
∧\ unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂)),
  &\\[2.5mm]
unshunt(\mysplit(φ₁))
  & \text{if }φ ≡ ¬ (φ₁ ⇒ φ₂)\\
∧\ unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
  &\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(¬ φ₂))
  & \text{if }φ ≡ ¬ (φ₁ ⇔ φ₂)\\
∧\ unshunt(¬ φ₂ ⇒ \mysplit(φ₁)),
  &\\[2.5mm]
unshunt(\mysplit(φ₁)),
  & \text{if }φ ≡ ¬ ¬ φ₁\\
⊤,
  & \text{if }φ ≡ ¬ ⊥\\
⊥,
  & \text{if }φ ≡ ¬ ⊤.
\end{cases}
\end{equation}

The definition above suffers termination problems since it is not
a structural recursion. Therefore, we have reformulated Def.~\ref{eq:split-definition} to solve such a issue following the technique
described in Subsection~\ref{ssec:structural-recursion}.
Here and below, we refer to $split$ as the function after applying
the process mentioned above.

\begin{lemma}[\thmsplit] $Γ ⊢ \mysplit(φ)$ then $Γ ⊢ φ$.
\label{lem:lem-split}
\end{lemma}

\begin{proof} We now proceed by induction on the structure of the
formula φ in the definition of Eq.~\ref{eq:split-definition}.
For a complete proof, we refer the reader to~\cite{AgdaMetis}.\\[3mm]

>>>>

\input{sections/diagram.tex}

% The proof-reconstruction approach proposed here consists of a series
% of steps similar to the workflow presented by \citeauthor{sultana2015} in
% \cite{sultana2015}. This process is a translation from a source
% system to a target system. In our case, the system of origin, the
% automatic theorem prover, is \Metis; the target system is a proof
% assistant, \Agda. We choose \Agda, but another proof assistant with the same support of type theory and inductive types is valid (see a description of \Agda for its main features in
% Section~\ref{sec:implementation}).
% The overview of the proof-reconstruction is presented in
% Fig.~\ref{fig:proof-reconstruction-workflow}, the process can begin with
% a problem in \CPL that we encode in \TPTP format to be sent as the
% input of the \Metis prover.
% The algorithms of \Metis in Step~3 find
% a solution if the problem is a theorem, in that case \Metis replies
% with a derivation in \TSTP format of the proof, otherwise the problem is counter feasible.

% With such a derivation, in Step~5, we process the derivation with our \Haskell translator tool, \Athena~\cite{Athena}.
% \Athena parses the \TSTP format, analyze the
% derivation and generates a representation of the natural deduction
% proof using a tree data structure (see the properties
% of this tree in Section~\ref{ssec:metis-proofs}).
% We perform an analysis of the proof-tree removing redundancies,
% unnecessary steps or unused output from the \TSTP derivation.

% As a result,
% \Athena generates an \Agda file with the proof-term. We have included
% in this file, \emph{imports} that call our \Agda libraries that give
% support for CPL presented in~\cite{AgdaProp}, and support for \Metis'
% inference rules implemented in~\cite{AgdaMetis}.

% Finally, we use the proof checker in Step~7 to type-check the
% proof-term. If type-checking succeed, the \TSTP
% derivation is correct, otherwise when it fails, one of the three
% actors in the process could be the responsible: \Metis, \Athena, or
% \Agda.

% We are responsible for errors committed in the translation from
% \TSTP to \Agda: bugs in the parsing module, lack of the coverage of
% \Metis' inference rules, and last but not least, the pretty print of
% the proof in \Agda code.

% In the rest of this section, we provide a formal description using
% type theory to build definitions and theorems for functions
% necessary to emulate \Metis' inference rules.
% Some conventions about notations. We will write the
% type judgments with ``$a : A$'' to denote that $a$ has type
% $\abbre{A}$. Types are written using small caps.
% For propositions, we will denote by $φ ≡ ψ$, the syntactic
% equivalence. We use the \Prop type for propositions and for theorems,
% the dependent type, $ Γ ⊢ φ$ where $Γ$ has type $\abbre{List} \Prop$ and
% $φ$ has type $\Prop$.

% % ...................................................................

% % section section_name (end)
% \subsection{Structural Recursion}
% \label{ssec:structural-recursion}

% To emulate \Metis' inference rules in type theory,
% we observed that transformations performed by some rules were given by \emph{general recursive} functions.
% But algorithms based on general recursions can not be translated
% directly in type theory since it is not a guarantee they terminate.

% Then, we solve the termination problem by modifying problematic functions
% to be \emph{structural recursive}~\cite{Coquand1992,Abel2002,Bove2005}.

% We use the same notion of \emph{structurally smaller} from
% \citeauthor{Abel2002}~\cite{Abel2002} to say, a recursive function is
% structurally recursive if it calls itself with only
% \emph{structurally smaller}
% arguments.

% To convert a general recursive call into a structurally recursive call,
% we use the \emph{Bounded Recursion} technique presented
% in~\cite{Bertot2004} but
% others found in \cite{Coquand1992,Abel2002,Bove2005} could be used.
% This technique converts a general recursive
% function into a structural recursive function by adding an additional
% argument for the function complexity, the bound for the number of times, the function needs to call itself to get the final result.

% Then, we add the natural number $n : \Nat$ in the input of the
% general recursive function, and perform the recursion over the structure of
% \Nat type. \abbre{Nat} type is the usual type for natural numbers with \zero and \suc as their constructors. We consider the common arithmetic operations ($+,\,-,\,*$) and numbers $0,1,2,\dots$ as representations of $\zero, \suc, \suc(\zero), \dots$.

% Then, the type-checker will
% recognize structurally smaller arguments by using the lexicographic
% order~\cite{Abel2002} for such kind of functions.

% In other words, we take a general recursive function $f: \Prop \to \Prop$.
% and we redefine that function as
% $f: \Prop \to \Nat \to \Prop$, and perform the recursion
% using induction on the structure of the \Nat type.

% Let us consider the following example to exhibit the technique
% mentioned above for an inner function in the \strip inference.

% \begin{align}
% \begin{split}
% \label{eq:unshunt-definition}
% &unshunt_{0}(φ) =\\
% &\hspace{3mm}\begin{cases}
% unshunt_{0}((φ₁ ∧ φ₂) ⇒ φ₃),
%   & \text{if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
% unshunt_{0}(φ₁ ⇒ φ₂) ∧ unshunt_{0}(φ₁ ⇒ φ₃),
%   & \text{if }φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
% φ, &\text{otherwise.}
% \end{cases}
% \end{split}
% \end{align}

% From the definition above, the size of the argument, ((φ₁ ∧ φ₂) ⇒ φ₃),
% is not structurally smaller than the input formula, (φ₁ ⇒ (φ₂ ⇒ φ₃)), since ((φ₁ ∧ φ₂) ⇒ φ₃) is not a subformula of (φ₁ ⇒ (φ₂ ⇒ φ₃)).

% Thus, we define the function
% $unshunt_{1} : \Prop \to \Nat \to \Prop$
% by structural recursion in its second argument in
% Eq.~(\ref{eq:unshunt-structured}).

% \begin{align}
% \label{eq:unshunt-structured}
% \begin{split}
% &unshunt_{1}(φ, \zero)\hspace{3mm}= φ\\
% &unshunt_{1}(φ, \suc(n)) =\\
% &\hspace{3mm}\begin{cases}
% unshunt_{1}((φ₁ ∧ φ₂) ⇒ φ₃, n),
%   &\text{if } φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
% unshunt_{1}(φ₁ ⇒ φ₂, n) ∧\ unshunt_{1}(φ₁ ⇒ φ₃, n),
%    &\text{if } φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃).
% \end{cases}
% \end{split}
% \end{align}

% As we mentioned above, the natural number in the second argument of the
% function used in Eq.~(\ref{eq:unshunt-structured}) stands for a complexity
% measure of the function, in this case $unshunt_{1}$ function.

% A \emph{complexity measure} defined in \cite{Agudelo-Agudelo2017}
% and presented in \cite{VanDalen1994} as \emph{the rank for a
% proposition}, is a function $\Prop \to \Nat$ that assigns a natural
% number for a proposition φ as follows.

% \begin{equation*}
% \label{eq:rank-definition}
% rank(φ)=
% \begin{cases}
% 0, &\text{for atomic }φ \\
% rank(φ₁) + c_{¬},  &\text{if } φ ≡ \neg φ₁ \\
% \max{(rank(φ₁),rank(φ₂))} + c_{\square},
%   &\text{if } φ ≡ φ₁\,\square\, φ₂.
% \end{cases}
% \end{equation*}

% where $c_{¬}$ or $c_{\square}$ is a non-negative integer called the
% complexity of the connective.

% Recall, we want to maintain the same outputs as the original unshunt
% does in Eq.~(\ref{eq:unshunt-definition}) and the complexity measure of it,
% is a bound for the number of times this function can call itself.

% Therefore, we define an appropriate complexity measure
% $unshunt_{cm}: \Prop → \NN$ for our $unshunt_{1}$ function presented in
% Eq.~(\ref{eq:unshunt-rank-definition}).

% \begin{align}
% \begin{split}
% \label{eq:unshunt-rank-definition}
% &unshunt_{cm}(φ) =\\
% &\hspace{3mm}
% \begin{cases}
% unshunt_{cm}(φ₃) + 2, &\text{if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
% \max(unshunt_{cm}(φ₂),unshunt_{cm}(φ₃)) + 1,
%   &\text{if }  φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
% 0, &\text{otherwise.}
% \end{cases}
% \end{split}
% \end{align}

% Finally, the curated version of $unshunt_{0}$ function is presented
% in Eq.~(\ref{eq:unshunt-fixed}).

% \begin{equation}
% \label{eq:unshunt-fixed}
% unshunt(\varphi) := unshunt_{1}(\varphi, unshunt_{cm}(\varphi)).
% \end{equation}

% For a more complete treatment of general recursion, and how to
% define well-founded and well-structural recursive functions, we refer
% the reader to Ref.~\cite{Bove2005}, where Bove and Capettra described
% not only general recursion, but they also treat mutual recursive
% function, among other kinds.

% % ................................................

% \subsection{Emulating \Metis' Inference Rules}
% \label{ssec:emulating-inferences}

% \subsubsection{Splitting a Goal.}
% \label{sssec:splitting-goal}

% The goal of a problem to be proved by \Metis is processed to generate new
% propositions, the new targets, \Metis calls these formulas
% \emph{subgoals}.

% The conjunction of these subgoals is a proposition equivalent to the
% original goal. Then, we find a refutation for each subgoal in the \Metis' \TSTP derivation.

% Nevertheless, \Metis does not provide neither an explanation for the user about the process in its \TSTP derivations nor in its documentation.

% In the following, we show how Theorem.~\ref{thm:thm-strip} proves the correctness of the process mentioned above. But first, let prove some lemmas for $unshunt$ function presented in Eq.~(\ref{eq:unshunt-fixed}),
% and $split$ function from the definition in Eq.~(\ref{eq:split-definition}).

% \begin{lemma}[lem-unshunt]
%   \label{lem:unshunt}
%   $Γ ⊢ unshunt(φ)$, then $Γ ⊢ φ$.
% \end{lemma}

% \begin{proof} Use induction on the cases defined by the outcome of the
% $unshunt$ function in Eq.~(\ref{eq:unshunt-structured}).
% When $n = 0$, by definition, $Γ ⊢ φ$.
% When $n = 1$, we apply the following theorem from~\cite{AgdaProp}

% \begin{equation*}
% \texttt{∧⇒-to-⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
% \end{equation*}

% For $n > 1$, we will use induction on the structure of second
% argument.

% (φ₁ ⇒ (φ₂ ⇒ φ₃))
% \begin{equation*}
%   \begin{bprooftree}
%   \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
%   \RightLabel{By def.}
%   \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
%   \RightLabel{lem-unshunt}
%   \UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
%   \RightLabel{∧⇒-to-⇒⇒}
%   \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).}
%   \end{bprooftree}
% \end{equation*}

% Second, using the following theorem from \cite{AgdaProp},

% \begin{equation*}
%   \texttt{⇒∧-to-⇒∧}\ :\ Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃) → Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃).
% \end{equation*}

% ($φ₁ ⇒ (φ₂ ∧ φ₃)$)
% \begin{equation*}
% \scalebox{0.95}{
% ($\mathcal{D}_1$)
% \begin{bprooftree}
% \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n) ∧ unshunt(φ₁ ⇒ φ₃, n)$}
% \RightLabel{∧-proj₁}
% \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n)$}
% \RightLabel{By def.}
% \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n-1)$}
% \RightLabel{lem-unshunt}
% \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$,}
% \end{bprooftree}
% }
% \end{equation*}

% \begin{equation*}
% (\mathcal{D}_2)
% \begin{bprooftree}
% \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂) ∧ unshunt(φ₁ ⇒ φ₃)$}
% \RightLabel{∧-proj₂}
% \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₃, n)$}
% \RightLabel{By def.}
% \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₃, n-1)$}
% \RightLabel{lem-unshunt}
% \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$,}
% \end{bprooftree}
% \end{equation*}

% \begin{equation*}
%   \begin{bprooftree}
%   \AxiomC{$\mathcal{D}_1$}
%   \AxiomC{$\mathcal{D}_2$}
%   \RightLabel{∧-intro}
%   \BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
%   \RightLabel{⇒∧⇒-to-⇒∧}
%   \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃).}
%   \end{bprooftree}
% \end{equation*}
% \end{proof}

% Now, let us consider the $split$ function presented in
% Eq.~(\ref{eq:split-definition}) that yields a conjunction of subgoals from a proposition, the goal in \Metis.

% \begin{equation}
% \label{eq:split-definition}
% \mysplit(φ) =
% \begin{cases}
% unshunt(\mysplit(φ₁)) &\text{if }φ ≡ φ₁ ∧ φ₂\\
% ∧\ unshunt(φ₁ ⇒ \mysplit(φ₂)), &\\[2.5mm]
% unshunt(¬ φ₁ ⇒ \mysplit(φ₂)),
%   & \text{if }φ ≡ φ₁ ∨ φ₂\\[2.5mm]
% unshunt(φ₁ ⇒ \mysplit(φ₂)),
%   & \text{if }φ ≡ φ₁ ⇒ φ₂\\[2.5mm]
% unshunt(φ₁ ⇒ \mysplit(φ₂))
%   & \text{if }φ ≡ φ₁ ⇔ φ₂\\
% ∧\ unshunt(φ₂ ⇒ \mysplit(φ₁)),
%   &\\[2.5mm]
% unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
%   & \text{if }φ ≡ ¬ (φ₁ ∧ φ₂)\\[2.5mm]
% unshunt(\mysplit(¬ φ₁))
%   & \text{if }φ ≡ ¬ (φ₁ ∨ φ₂)\\
% ∧\ unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂)),
%   &\\[2.5mm]
% unshunt(\mysplit(φ₁))
%   & \text{if }φ ≡ ¬ (φ₁ ⇒ φ₂)\\
% ∧\ unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
%   &\\[2.5mm]
% unshunt(φ₁ ⇒ \mysplit(¬ φ₂))
%   & \text{if }φ ≡ ¬ (φ₁ ⇔ φ₂)\\
% ∧\ unshunt(¬ φ₂ ⇒ \mysplit(φ₁)),
%   &\\[2.5mm]
% unshunt(\mysplit(φ₁)),
%   & \text{if }φ ≡ ¬ ¬ φ₁\\
% ⊤,
%   & \text{if }φ ≡ ¬ ⊥\\
% ⊥,
%   & \text{if }φ ≡ ¬ ⊤.
% \end{cases}
% \end{equation}

% The definition above suffers termination problems since it is not
% a structural recursion. Therefore, we have reformulated Def.~\ref{eq:split-definition} to solve such a issue following the technique
% described in Section~\ref{ssec:structural-recursion}.
% Here and below, we refer to $split$ as the function after applying
% the process mentioned above.

% \begin{lemma}[\thmsplit] $Γ ⊢ \mysplit(φ)$ then $Γ ⊢ φ$.
% \label{lem:lem-split}
% \end{lemma}

% \begin{proof} We now proceed by induction on the structure of the
% formula φ in the definition of Eq.~(\ref{eq:split-definition}).
% For a complete proof, we refer the reader to~\cite{AgdaMetis}.\\[3mm]

% ($φ ≡ φ₁ ∧ φ₂$). We get a proof for each conjunct and using the introduction rule for conjunction, the result follows.

% \begin{equation*}
% (\mathcal{D})\hspace{3mm}
%   \begin{bprooftree}
%   \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
%   \RightLabel{By def.}
%   \UnaryInfC{$Γ ⊢ unshunt(φ₁) ∧ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
%   \RightLabel{∧-proj₁}
%   \UnaryInfC{$Γ ⊢ unshunt(φ₁)$}
%   \RightLabel{lem-unshunt}
%   \UnaryInfC{$Γ ⊢ φ₁$.}
%   \end{bprooftree}
% \end{equation*}
